---
title: 'Chapter 3: Sampling the Imaginary'
output:
  github_document
---
```{r, setup, warning=FALSE, message=FALSE}
library(rethinking)
library(magrittr)
data(homeworkch3)
sessionInfo()
```

## Easy
Same setup as the globe-tossing example in the chapter
```{r, easy_setup}
grid_size <- 1000
num_samples <- 1e4
p_grid <- seq(from=0, to=1, length.out=grid_size)
prior <- rep(1, grid_size)

binom_likelihood <- function(prob, trials, successes) {
  # Return vector of likelihood weights corresponding to
  # success probability `prob` given a fixed number of `trials`
  # and `successes`
  return(dbinom(x=successes, size=trials, prob=prob))
}

likelihood <- binom_likelihood(p_grid, 9, 6)  ## example in text
posterior <- prior * likelihood
posterior <- posterior / sum(posterior)

set.seed(100)
samples <- sample(p_grid, prob=posterior, size=num_samples, replace=TRUE)
hist(samples, breaks=50)
```

### 3E1
How much posterior probability lies below $p = 0.2$?
```{r, 3e1}
mean(samples < 0.2)
```

### 3E2
How much posterior probability lies above $p = 0.8$?
```{r, 3e2}
mean(samples > 0.8)
```

### 3E3
How much posterior probability lies between $p = 0.2$ and $p = 0.8$?
```{r, 3e3}
mean(samples >= 0.2 & samples <= 0.8)
```

### 3E4
20% of the posterior probability lies below which value of $p$?
```{r, 3e4}
quantile(samples, probs=0.2)
```

### 3E5
20% of the posterior probability lies above which value of $p$?
```{r, 3e5}
quantile(samples, probs=0.8)
```

### 3E6
What is the narrowest 66% interval for the posterior probability?
```{r, 3e6}
HPDI(samples, prob=0.66)
```

###  3E7
What is the 66% interval that has the same probability above and below it?
```{r, 3e7}
PI(samples, prob=0.66)  ## using the function in rethinking
prob_tail = (1 - 0.66) / 2
quantile(samples, probs=c(prob_tail, 1-prob_tail)) ## using base functions
```

## Medium
### 3M1
Suppose the sequence was 8 water in 15 tosses. Construct the posterior starting
from the same prior.
```{r, 3m1}
likelihood2 = binom_likelihood(prob=p_grid, trials=15, successes=8)
posterior2 = prior * likelihood2
posterior2 = posterior2 / sum(posterior2)
plot(p_grid, posterior2, type='l', lwd=2)
abline(v=0.5, lty=3)
```

### 3M2
Draw `r num_samples` samples from this new distribution. What is the 90% HPDI for p?
```{r, 3m2}
samples2 = sample(p_grid, size=num_samples, prob=posterior2, replace=TRUE)
HPDI(samples2, .9)
```

### 3M3
Construct a posterior predictive check from this model. What is the probability
of observing 8 water in 15 tosses?
```{r, 3m3}
water_in_15 = rbinom(n=num_samples, size=15, prob=samples2)
mean(water_in_15 == 8)
water_in_15 %>% table %>% barplot
```

### 3M4
Keeping the posterior from the 8 out of 15 data, calculate the probabilty of
observing 6 water in 9 tosses.
```{r, 3m4}
water_in_9 = rbinom(n=num_samples, size=9, prob=samples2)
mean(water_in_9 == 6)
water_in_9 %>% table %>% barplot
```

### 3M5
Repeat 3M1 - 3M4 starting from a prior that is 0 below $p = 0.5$ and constant
otherwise.
```{r, 3m5}
# define new prior
prior3 = ifelse(p_grid < 0.5, 0, 1)
rle(prior3)  ## checking, should be two chunks of 500

# get new posterior
posterior3 = prior3 * likelihood2
posterior3 = posterior3 / sum(posterior3)
plot(p_grid, posterior3, type='l', lwd=2, col='orange')

# get samples of p and the 90% highest posterior density interval
samples3 = sample(p_grid, size=num_samples, prob=posterior3, replace=TRUE)
HPDI(samples3, prob=0.9)
hist(samples3, breaks=50)

# what's the chance of getting 8 water in 15 tosses from generated data?
water_in_15_new = rbinom(n=num_samples, size=15, prob=samples3)
mean(water_in_15_new == 8)
water_in_15_new %>% table %>% barplot

# what's the chance of getting 6 water in 9 using the same distro for p?
water_in_9_new = rbinom(n=num_samples, size=9, prob=samples3)
mean(water_in_9_new == 6)
water_in_9_new %>% table %>% barplot
```

## Hard
All problems use the same data available in the book package. The data shows the
gender for the first child `birth1` and the second child `birth2` for 100
2-child families (male: 1, female: 0).

### 3H1
```{r, 3h1}
prior4 <- rep(1, grid_size)
all_births = c(birth1, birth2)
likelihood4 = binom_likelihood(prob=p_grid,
                               trials=length(all_births),
                               successes=sum(all_births))
posterior4 = prior4 * likelihood4
posterior4 = posterior4/sum(posterior4)
(p_max = p_grid[which.max(posterior4)])   ## define and return
plot(p_grid, posterior4, type='h', col='grey')
abline(v=p_max, lty=3)
```

### 3H2
Draw `r num_samples` samples from the posterior above. Get the 50%, 89%, and 97%
highest density posterior intervals.
```{r, 3h2}
samples4 = sample(p_grid, size=num_samples, prob=posterior4, replace=TRUE)
HPDI(samples=samples4, 0.5)
HPDI(samples=samples4, 0.89)
HPDI(samples=samples4, 0.97)
```

### 3H3
Simulate `r num_samples` of male births in 200 childbirths based on the current
model. Show the distribution of this number and compare it to the outcome in the
data of `r sum(all_births)` males.
```{r, 3h3}
males_in_200 = rbinom(num_samples, size=200, prob=samples4)
males_in_200 %>% density %>% plot(main='Simulated number of male births in 200')
abline(v=sum(all_births), lty=3)
```

### 3H4
Do the same thing for the 100 first-born births.
```{r, 3h4}
males_in_100 = rbinom(num_samples, size=100, prob=samples4)
males_in_100 %>% density %>% plot(main='Simulated number of male births in 100')
abline(v=sum(birth1), lty=3, col='orange')
```

### 3H5
Focus on the births of those boys who followed girls. Find out how many times
this happened (number of attempts) and how often these were boys (number of 
successes, in this notation).
```{r, 3h5}
# get basic data
first_girl = birth1 == 0
birth_after_girl = birth2[first_girl]
(trials_after_girl = length(birth_after_girl))
(num_boy_after_girl = sum(birth_after_girl))

# simulate
males_after_girl = rbinom(num_samples, size=trials_after_girl, prob=samples4)
males_after_girl %>% density %>% plot(main=paste('Number of male births in', trials_after_girl))
abline(v=num_boy_after_girl, lty=3, col='purple')
```
