---
title: 'Chapter 4: Linear Models'
output:
  #github_document
  md_document:
    #variant: gfm
    pandoc_args: --webtex
    
---
```{r, setup, warning=FALSE, message=FALSE}
library(rethinking)
library(magrittr)
set.seed(2020)
sessionInfo()

data(Howell1)
```

## Medium

### 4M1
```{r, 4m1}
nsamples = 1e4
prior_sample_m1 = rnorm(n=nsamples,
                        mean=rnorm(nsamples, 0, 10),
                        sd=runif(nsamples, 0, 10))
prior_sample_m1 %>% hist(breaks=50)
```

### 4M2
```{r, 4m2}
model_m2 <- alist(
  y ~ dnorm(mu, sigma),
  mu ~ dnorm(0, 10),
  sigma ~ dunif(0, 10)
)
```

### 4M3
\[
  \begin{align*}
    y_i &\sim \text{Normal}(\mu_i, \sigma) \\
    \mu_i &= \alpha + \beta x_i \\
    \alpha &\sim \text{Normal}(0, 50) \\
    \beta &\sim \text{Uniform}(0, 10) \\
    \sigma &\sim \text{Uniform}(0, 50)
  \end{align*}
\]

### 4M4
All units are in centimeters. I'm assuming "students" means university students.

\[
  \begin{align*}
    y_i &\sim \text{Normal}(\mu_i, \sigma) \\
    \mu_i &= \alpha + \beta x_i \\
    \alpha &= \text{Normal}(160, 20) \\
    \beta &= \text\{Uniform}(0, 20) \\
    \sigma &= \text{Uniform}(0, 20)
  \end{align*}
\]

### 4M5
I already assumed students get taller every year ($\beta$ cannot be negative). 
Based on the new information, my prior for $\alpha$ is now $\alpha \sim \text{Normal}(120, 10)$.
I've narrowed the standard deviation, because 20 seems excessive given the mean
is 120. It would imply there is a reasonable chance we have students of of 80cm.

### 4M6
That means the standard deviation $\sigma$ is never higher than 8cm. So it's
prior now becomes: $\sigma \sim \text{Uniform}(0, 8)$.

## Hard
### 4H1
Let's refit a simple model, and then produce estimates for these points.
```{r, 4h1}
dd <- Howell1
w_mean <- mean(dd$weight)
w_sd <- sd(dd$weight)
dd$weight_c <- (dd$weight - w_mean) / w_sd
dd$weight_c2 <- dd$weight_c^2

m4h1 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * weight_c + c * weight_c2,
    a ~ dnorm(168, 100),
    b ~ dnorm(0, 20),
    c ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), data=dd)
m4h1

weight_x <- c(46.95, 43.72, 64.78, 32.59, 54.63)
weight_xc <- (weight_x - w_mean) / w_sd
weight_xc2 <- weight_xc^2
pred_4h1 <- sim(m4h1, data=data.frame(weight_c=weight_xc, weight_c2=weight_xc2))
cbind(weight_x, colMeans(pred_4h1), t(apply(pred_4h1, 2, HPDI)))
```

### 4H2
```{r, 4h2}
dd2 <- Howell1[Howell1$age < 18, ]
str(dd2)

w_mean2 <- mean(dd2$weight)
w_sd2 <- sd(dd2$weight)

dd2$weight_c <- (dd2$weight - w_mean2) / w_sd2
dd2$weight_c2 <- dd2$weight_c^2

m4h2 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight_c + c*weight_c2,
    a ~ dnorm(120, 80),
    b ~ dnorm(0, 30),
    c ~ dnorm(0, 30),
    sigma ~ dunif(0, 50)
  ), data=dd2)
m4h2

plot(height ~ weight, data=dd2, pch=20)

w_range <- range(dd2$weight)
w_x <- seq(from=w_range[1], to=w_range[2], length.out=100)
w_x_c <- (w_x - w_mean2) / w_sd2
w_dta <- data.frame(weight_c=w_x_c, weight_c2=w_x_c^2)
m4h2_link <- link(m4h2, data=w_dta)
m4h2_sim <- sim(m4h2, data=w_dta)
shade(apply(m4h2_link, 2, HPDI), w_x)
shade(apply(m4h2_sim, 2, HPDI), w_x)
```

The centered quadratic fit is not a bad one. A linear model (my first intuition)
would not have been a good idea.

### 4H3
```{r, 4h3}
m4h3 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*log(weight),
    a ~ dnorm(178, 100),
    b ~ dnorm(0, 100),
    sigma ~ dunif(0, 50)
  ), data=dd2)
m4h3

plot(height ~ weight, data=dd2, pch=20)

w_dta2 <- data.frame(weight=w_x)
m4h3_link_interval = apply(link(m4h3, data=w_dta2), 2, HPDI)
m4h3_sim_interval = apply(sim(m4h3, data=w_dta2), 2, HPDI)
shade(m4h3_link_interval, w_x)
shade(m4h3_sim_interval, w_x)
```
The log fit is about as good as the quadratic model, but with much less feature
modifications.